---
title: "RWorksheet_Regacho#3b"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Create a data frame using the table below.

a. Write the codes.
```{r}
ID <- 1:6
Sex <- c("Male", "Female", "Male", "Female", "Female", "Male")
Type_of_House <- c("Wood", "Concrete", "Semi-Concrete", "Wood", "Concrete", "Wood")
Father_Occupation <- c("Farmer","Driver", "Other", "Farmer", "Driver", "Farmer")
Siblings_Attending <- c(3,5,2,6,5,4)

household <- data.frame(ID,Sex, Type_of_House, Father_Occupation, Siblings_Attending,
stringsAsFactors = FALSE)
household
```

b. Describe the data. Get the structure or the summary of the data
```{r}
str(household)
```


c. Is the mean number of siblings attending is 5?
```{r}
mean_siblings <-
mean(household$Siblings_Attending)
mean_siblings
mean_siblings == 5

```
d. Extract the 1st two rows and then all the columns using the subsetting functions. Write the codes and its output.
```{r}
household[1:2,]
```
e. Extract 3rd and 5th row with 2nd and 4th column. Write the codes and its result.
```{r}
household[c(3,5), c(2,4)]
```
f. Select the variable types of houses then store the vector that results as types_houses.
Write the codes.
```{r}
types_houses <- household$Type_of_House
types_houses
```
g. Select only all Males respondent that their father occupation was farmer. Write the codes and its output.
```{r}
male_farmers <- subset(household, Sex == "Male" & Father_Occupation == "Farmer")
male_farmers
#subset() is readable and convenient
```
h. Select only all females respondent that have greater than or equal to 5 number of siblings attending school. Write the codes and its outputs.
```{r}
female_ge5 <- subset(household, Sex == "Female" & Siblings_Attending >=5)
female_ge5
```


2. Write a R program to create an empty data frame. Using the following codes:
df = data.frame(Ints=integer(),

Doubles=double(), Characters=character(),
Logicals=logical(),
Factors=factor(),
stringsAsFactors=FALSE)

print("Structure of the empty dataframe:")
print(str(df))
a. Describe the results.
```{r}
df= data.frame(Ints=integer(), Doubles= double(), Characters= character(), Logicals= logical(), Factors= factor(), stringsAsFactors = FALSE)
print("Structure of the empty dataframes:")
str(df)
```
3. Create a .csv file of this. Save it as HouseholdData.csv
```{r}
write.csv(household, file = "HouseholdData.csv", row.names = FALSE)
```
Figure 2: Figure 2: Sentiment Analysis
a. Import the csv file into the R environment. Write the codes.
```{r}
household_imported <- read.csv("HouseholdData.csv", stringsAsFactors = FALSE)
head(household_imported)
# Save then import to demonstrate persistence/reproducibility. In RStudio make sure your working directory is set to the folder you want to use.
```
b. Convert the Sex into factor using factor() function and change it into integer.[Legend:Male = 1 and Female = 2]. Write the R codes and its output.
```{r}
household_imported$Sex <-
factor(household_imported$Sex, levels = c("Male", "Female"))
as_integer_sex <-
as.integer(household_imported$Sex)
data.frame(Sex = household_imported$Sex, Sex_int = as_integer_sex)
#factor(..., levels =..) sets the internal codes - Male to 1, Female to 2.
```
c. Convert the Type of Houses into factor and change it into integer. [Legend: Wood= 1; Congrete = 2; Semi-Congrete = 3]. Write the R codes and its output.
```{r}
household_imported$Type_of_House <-
factor(household_imported$Type_of_House, levels = c("Wood", "Concrete", "Semi-Concrete"))
household_imported$Type_of_House_int <-
as.integer(household_imported$Type_of_House)
household_imported
```


d. On father’s occupation, factor it as Farmer = 1; Driver = 2; and Others = 3. What is the R code and its output?
```{r}
household_imported$Father_Occupation <-
  as.integer(household_imported$Father_Occupation)
household_imported
```
e. Select only all females respondent that has a father whose occupation is driver. Write the codes and its output.
```{r}
female_driver <- subset(household_imported, Sex == "Female" & Father_Occupation == "Driver")
female_driver
```
f. Select the respondents that have greater than or equal to 5 number of siblings attending school. Write the codes and its output.
```{r}
sib_ge5 <- subset(household_imported, Siblings_Attending >=5)
sib_ge5
```
4. Interpret the graph.

#The bar plot shows that Concrete houses are the most common, followed by Wood. The Farmer occupation appears most often among fathers, and households with 5 or more siblings attending school are less frequent. This suggests a concentration of certain house types and occupations in the sample and a skew toward smaller numbers of attending siblings

